---
title: "caretEnsemble Setup"
author: "Miguel Conde"
date: "21 de marzo de 2016"
output: html_document
---

```{r me, echo = FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE)
```

# Load `caretEnsemble` and allow Parallel Processing
```{r CARETENSEMBLE_AND_PARALLEL}
library(caretEnsemble)

library(doParallel)
cl <- makeCluster(max(detectCores()-1,1))
registerDoParallel(cl)
```

Loading data

```{r LOAD_DATA}
library(C50)
data(churn)
churn <- rbind(churnTrain, churnTest)

str(churn)
```

Creating dummy vars (for xgboost)
```{r DUMMY_VARS}
library(caret)
dummies <- dummyVars(churn ~ ., data = churn)
newChurn <- as.data.frame(predict(dummies, newdata = churn))

cols_no <- grep(".*no", names(newChurn))
newChurn <- newChurn[, setdiff(1:ncol(newChurn), cols_no)]

```

Training and test sets

```{r TRAIN_AND_TEST_SETS}
set.seed(1234)
trainingIdx <- createDataPartition(churn$churn, list = FALSE)

trainPred <- newChurn[trainingIdx, ]
trainClass <- churn[trainingIdx, "churn"]

testPred <- newChurn[-trainingIdx, ]
testClass <- churn[-trainingIdx, "churn"]
```

## Bunch of models with `caretEnsemble`
Bunchs of models are easy with `caretEnsemble`:
```{r CARET_LIST, cache = TRUE}
# Example of Stacking algorithms
# create submodels
library(caretEnsemble)

classModels <- c("ada","AdaBag","AdaBoost.M1","avNNet","Boruta","C5.0",
                 "earth","gbm", "knn","lda","nb","nnet","ranger","svmLinear",
                 "svmLinear2", "svmPoly", "svmRadial","xgbLinear", "xgbTree")

trCtrl <- trainControl(method            = "cv", # Need more than 1
                       number            = 10,   # for later resampling
                       verboseIter       = FALSE,
                       classProbs        = TRUE)

set.seed(5678)
modelsList <- caretList(x          = trainPred,
                        y          = trainClass, 
                        trControl  = trCtrl, 
                        methodList = classModels)
```


Compare models
```{r COMPARE_CARET_LIST}
resamps <- resamples(modelsList)
summary(resamps)

xyplot(resamps, what = "BlandAltman")
dotplot(resamps)
densityplot(resamps)
bwplot(resamps)
splom(resamps)
parallelplot(resamps)

diffs <- diff(resamps)
summary(diffs)

sort(resamps, decreasing = TRUE, metric = "Kappa")
```

Don't know, but these results don't look like those above, where i made a bunch of models by hand.

Predicting with `caretEnsemble`
```{r PREDICT_CARET_ENSEMBLE}
p <- as.data.frame(predict(modelsList[], newdata=head(testPred)))
print(p)

p <- as.data.frame(predict(modelsList[], newdata=head(testPred), type = "prob"))
print(p)
```

### Understanding `caretList()`

`caretList()` is a flexible function for fitting many different caret models, 
with the same resampling parameters, to the same dataset. It returns a 
convenient list of caret objects which can later be passed to caretEnsemble 
and caretStack. `caretList()` has almost exactly the same arguments as `train()` 
(from the `caret` package), with the exception that the `trControl` argument 
comes last. It can handle both the formula interface and the explicit `x`, `y` 
interface to train. As in caret, the formula interface introduces some 
overhead and the `x`, `y` interface is preferred.
 
`caretList()` is used to build lists of caret models on the same training data, 
with the same re-sampling parameters.

If you desire more control over the model fit, use the `caretModelSpec()` function to 
contruct a list of model specifications for the `tuneList` argument. This 
argument can be used to fit several different variants of the same model, 
and can also be used to pass arguments through train down to the component 
functions (e.g. trace=FALSE for nnet)

Notice that we are explicitly setting the resampling index to being used in 
`trainControl`. If you do not set this index manually, caretList will attempt to 
set it for automatically, but it“s generally a good idea to set it yourself:
```{r CARET_LIST_1}
trCtrl <- trainControl(method            = "cv", # Need more than 1
                       number            = 10,   # for later resampling
                       verboseIter       = FALSE,
                       savePredictions   = "final",
                       classProbs        = TRUE,
                       index             = createResample(trainClass, 10))
```

Models for the methodList argument.
`methodList` is a simple character vector of methods that will be fit with the 
default train parameters.
```{r CARET_LIST_2}
classModels <- c("ranger","xgbLinear") 
```

Model specifications for the `tuneList` argument, using `caretModelSpec()`
```{r CARET_LIST_3}
tuneModels  <- list(
  rf1 = caretModelSpec(method     = "rf", 
                       tuneGrid   = data.frame(.mtry=2)
                       ),
  rf2 = caretModelSpec(method     = "rf", 
                       tuneGrid   = data.frame(.mtry=10), 
                       preProcess ="range"
                       ),
  nn  = caretModelSpec(method     = "nnet",
                       preProcess = c("center", "scale"),
                       tuneLength = 2, 
                       trace      = FALSE
                       )
)
```

Here we go with `caretList()`
We will ask to build 5 models, two of them -"ranger","xgbLinear"- with the 
`train()` default arguments and three - "rf1", "rf2", "nn" - with specific 
`train()` arguments
```{r  CARET_LIST_4, cache = TRUE}
newModelsList <- caretList(# Arguments to pass to train() as '...' ###
                           x          = trainPred,
                           y          = trainClass, 
                           metric     = "Kappa",
                           ###########################################
                           # caretList() specific arguments ##########
                           trControl  = trCtrl,
                           methodList = classModels,
                           tuneList   = tuneModels
                           ###########################################
                           )
```

Predicting
```{r}
p <- as.data.frame(predict(newModelsList[],
                           newdata = head(testPred)))
print(p)
p <- as.data.frame(predict(newModelsList[], 
                           newdata = head(testPred), type = "prob"))
print(p)
```


## MODELS ENSEMBLES

[KAGGLE ENSEMBLING GUIDE](http://mlwave.com/kaggle-ensembling-guide/)

Should use low correlated models
```{r CORR_MODELS}
# ?modelCor

# modelCor(resamps)

bestModels <- sort(resamples(modelsList), 
                   decreasing = TRUE, 
                   metric = "Kappa")[1:5]
modelCor(resamples(modelsList[bestModels]))
```

Correlation doesn't resemble that of bunch of models...

### Ensembles

#### Voting
TBD

#### Averaging
TBD

#### Weighted voting and averaging
TBD

### Stacking 

```{r STACK_GLM, cache = TRUE}
# stack using gbm
set.seed(910)

bestModelsList <- modelsList[bestModels]
class(bestModelsList) <- class(modelsList)

stack.glm <- caretStack(
  bestModelsList, 
  method    = "glm", 
  metric    = "Kappa", 
  trControl = trainControl(method          = "cv",
                           number          =  10,
                           savePredictions = "final",
                           classProbs      = TRUE,
                           index           = createResample(trainClass, 10),
                           summaryFunction = twoClassSummary)
  )

save(stack.glm,      file = "./DATA/stack.glm.CE.Rda")
save(modelsList,     file = "./DATA/modelsList.CE.Rda")
save(bestModelsList, file = "./DATA/bestModelsList.CE.Rda")
save(newModelsList,  file = "./DATA/newModelsList.CE.Rda")

print(stack.glm)
summary(stack.glm)
```
**BUT IT CAN STACK ONLY 2 CATEGORIES CLASSIFIERS**

(In [A Brief Introduction to caretEnsemble](https://cran.r-project.org/web/packages/caretEnsemble/vignettes/caretEnsemble-intro.html) it's said *"DO NOT use the trainControl object you used to fit the training models to fit the ensemble"*)

```{r}

confusionMatrix(predict(stack.glm, newdata = testPred), testClass)

# varImp(stack.xgb)

library("caTools")
model_preds <- lapply(bestModelsList, 
                     predict, 
                     newdata = testPred, 
                     type    = "prob")
model_preds <- lapply(model_preds, function(x) x[,"yes"])
model_preds <- data.frame(model_preds)
ens_preds   <- predict(stack.glm, newdata = testPred, type = "prob")
model_preds$ensemble <- ens_preds
head(model_preds)
# CF <- coef(stack.xgb$ens_model$finalModel)[-1]
caTools::colAUC(model_preds, testClass)
# CF/sum(CF)

```

```{r}

model_preds <- lapply(bestModelsList, 
                     predict, 
                     newdata = testPred)
# model_preds <- lapply(model_preds, function(x) x[,"yes"])
model_preds <- data.frame(model_preds)
ens_preds   <- predict(stack.glm, newdata = testPred)
model_preds$ensemble <- ens_preds

head(model_preds)

```

I was afraid something like this was going to happen...

Predicting
```{r}
p <- as.data.frame(predict(stack.glm, newdata = head(testPred)))
print(p)
p <- as.data.frame(predict(stack.glm, newdata = head(testPred), type = "prob"))
print(p)
```

```{r}
# plot(stack.glm)
# dotplot(stack.glm)
```

Stop Parallel Processing
```{r STOP_PARALLEL}
stopCluster(cl)
```